Search for Spatially Extended \fermi-LAT Sources Using Two Years of Flight Data

### 2011-08-19 extended_catalog/paper/extended_catalog.tex
eps

% for highlighting text
\usepackage{framed}
\colorlet{shadecolor}{yellow}

@@

% some units
\newcommand{\kev}{\text{Kev}\xspace}
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\phflux}{\ensuremath{\ph\ \cm^{-2}\s^{-1}}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_{\text{ext}}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_{\text{inc}}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}
\newcommand{\likelihood}{\ensuremath{L}\xspace}
\newcommand{\chandra}{\text{{\em Chandra}}\xspace}
\newcommand{\swiftxrt}{\text{{\em Swift}/XRT}\xspace}
\newcommand{\rosat}{\text{{\em ROSAT}}\xspace}
\newcommand{\suzaku}{\text{{\em Suzaku}}\xspace}
\newcommand{\asca}{\text{{\em ASCA}}\xspace}
\newcommand{\einstein}{\text{{\em Einstein}}\xspace}
\newcommand{\xmmnewton}{\text{{\em XMM-Newton}}\xspace}

@@

\renewcommand{\deg}{\ensuremath{^\circ}\xspace}

% the program names
\newcommand{\pointlike}{\text{\em pointlike}\xspace}
\newcommand{\python}{\text{\em python}\xspace}
\newcommand{\gtlike}{\text{\em gtlike}\xspace}
\newcommand{\gtobssim}{\text{\em gtobssim}\xspace}
\newcommand{\minuit}{\text{\em Minuit}\xspace}
\renewcommand{\approx}{\sim\!\xspace}

\shorttitle{Search for Extended LAT Sources}

@@

\title{Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight
Data}

@@

\begin{abstract}
We present a new method for analyzing 
the spatial extension of sources with
the Large Area Telescope (LAT), the primary science instrument on the
{\em Fermi Gamma-ray Space Telescope} (FGST). 
Spatial extension
is an important input to correctly 
associate LAT sources with their counterparts at other
wavelengths. We perform
a series of Monte Carlo studies to validate this tool and calculate
the LAT's detection threshold to detect the extension of 
spatially extended sources.  We then
apply this tool to test all sources in the two year source catalog for
extension. We report the detection of nine spatially extended sources
in addition to the twelve spatially extended sources reported in the
second Fermi-LAT Catalog (2FGL).
\end{abstract}

\listoftodos

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the Fermi
Gamma-Ray Space Telescope (FGST) that been surveying the $\gamma$-ray
sky since June 2008.  The LAT has a broad energy coverage (20 \mev to $>300$
\gev), wide field of view ($\approx 2.4 \sr$), and large effective area
($\approx 8000 \cm^2$ at $>1 \gev$).

@@

For each of these sources, the extension was determined based on
dedicated analysis of the particular source 
instead of being the outcome of a systematic search of the LAT catalog.
For the systematic scan of all LAT catalog sources presented in this work, we
therefore expect to find additional spatially extended LAT sources, in
particular since many extended sources in the Galactic plane have been
detected at \tev energies using air Cherenkov detectors and \tev and
\gev emission often originates from the same population of high-energy
particles.  A variety of spatially extended sources have been detected
in the Galactic plane at \tev energies. Most prominent was a survey of
the Galactic plane using the High Energy Stereoscopic System (H.E.S.S)
which reported 14 spatially extended sources
 with extensions varying from
$~0.1\deg$ to $~0.25\deg$ \citep{hess_plane_survey}.
In fact, within
our Galaxy only very few sources (most notably the $\gamma$-ray binaries
LS\,5039 \citep{HESSLS5039}, LS I+61-303 \citep{MAGICLSI, VERITASLSI}
HESS\,J0632+057 \citep{HESS0632}, and the Crab nebula \citep{crab_weekes}
have 
no significant extension detected at \tev energies. Given
these findings, it can be expected that if the source population at \gev
energies has similar properties, that at least some of the Fermi-LAT
sources in our Galaxy should be spatially resolvable at \gev energies.

\begin{shaded} 
High energy $\gamma$-rays are produced by the decay of $\pi^0$s produced
by hadronic interactions with interstellar matter and by relativistic
electrons due to Inverse Compton (IC) scattering and Bremsstrahlung
radiation.  Many of these spatially extended \tev sources in the
Galactic plane are unidentified. Some of these sources could produce
hadronic cosmic rays up to $\approx 10^{15}$ eV and subsequently radiate
$\gamma$-rays through $\pi^0$ decay \citep{blandford_and_eichler_1987}.
Studying the spectra of these \tev sources at \gev energies may help to
determine the emission mechanisms of these high energy sources.
\end{shaded}

@@

Morphological studies of sources in the \gev energy range using the
LAT are challenging because of the strongly energy-dependent PSF.
Additional complications arise for sources along the Galactic plane due to
systematic uncertainties in the Galactic diffuse emission.  The LAT's
PSF is limited at its lowest detectable energies by multiple scattering in the tracker
and the 68\% containment radius of the PSF is approximately 5.1\deg
at 100 \mev  when averaged over instrument acceptance and including
photons which convert in either the thick or thin layers of the
tracker. The PSF improves with energy and approaches a limit given
by the granularity of the silicon strips and is 0.14\deg at 10 \gev
\citep{on_orbit_calibration}.  However, since most high energy
astrophysical sources have power-law spectra, the better resolution of the higher
energy photons is offset by their low
statistics. Therefore a detailed investigation has been performed in this
work to determine the optimal energy range to measure source extension
(section \ref{extended_source_search_method}).

@@

We estimate the error on the extension of a source by fixing the position
of the source and varying the extension until the likelihood has fallen
by $\tfrac{1}{2}$, corresponding to a $1\sigma$ error.  This method is
shown schematically in figure~\ref{four_plots_ic443}(a) which shows the
change in \loglikelihood when varying the extension of the SNR IC443.
The localization error is calculated by fixing the extension and spectrum
of the source and fitting to the likelihood function a Gaussian as a function
of position.

@@

We tested two different models for the
surface brightness profile, either a two dimensional Gaussian
\begin{equation}\label{gauss_pdf}
  I(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform disk
\begin{equation}\label{disk_pdf}
  I(x,y)=
  \begin{cases}
    \frac{1}{\pi\sigma^2} & x^2+y^2\le\sigma^2 \\
    0                      & x^2+y^2>\sigma^2
  \end{cases}
\end{equation}
Although these shapes are significantly different, their convolution
with the PSF look similar.  Figure~\ref{compare_disk_gauss}
demonstrates this by showing $I_\text{src}$, the PSF, and the PDF for
a radially symmetric uniform brightness and the Gaussian surface brightness of size 0.5\deg.\footnote{To
allow a valid comparison between Gaussian and disk shaped morphologies,
we define the source size for this test as the radius containing
68\% of the intensity ($\rsixeight$).  $\rsixeight_\text{,Gaussian}=1.51\sigma$
and $\rsixeight_\text{,disk}=0.82\sigma$ where $\sigma$
is defined in equation~\ref{gauss_pdf} and
equation~\ref{disk_pdf} respectively.} This plot shows that the LAT is 
not very sensitive to probing the exact structure of an extended source.
In the search for extended sources, we therefore use only a uniform
disk as our spatial extension shape. We then quote the radius of the
disk edge as the size of the source.

@@

We test the performance of \pointlike fitting the extension of
point-like sources.
\cite{mattox_egret} discuss that the test statistic distribution
for a likelihood ratio test on the existence of a source at
a given position is 
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\tfrac{1}{2}(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of equation \ref{ts_ext_distribution} is
due to the null hypothesis (source flux $\Phi=0$) residing
on the edge of parameter space and the model hypothesis
adding a single degree of freedom. It is plausible to
expect a similar distribution of the test statistic
in the test for source extension as the same conditions
apply (with the source flux $\Phi$ replaced by the source radius $r$ and
$r<0$ being unphysical).
To validate this claim, we performed a Monte Carlo
study to calculate empirical distributions for $\tsext$ and
compared them to equation~\ref{ts_ext_distribution}.

@@

For each set of spectral parameters, $\approx30,000$ statistically independent
simulations were performed. For the dimmer spectral models, many of the
simulations left the source undetected ($\ts<25$)
and were discarded.  Table~\ref{ts_ext_num_sims}
shows the different spectral models used in our study as well as the
number of simulations.  The cumulative density of \tsext is plotted in
figure~\ref{ts_ext_mc}. The $\chi^2_1/2$ distribution of
equation~\ref{ts_ext_distribution} is overlaid for comparison.

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The LAT's flux threshold for a significant detection
of source extension drops quickly with
increasing source size and reaches a minimum around 0.5\deg. 
Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only 1-100\gev photons.
Except for very large ($>1\deg$) sources, our detection threshold is
not substantially improved including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in figure~\ref{four_plots_ic443}(b)
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\approx2.4$, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we therefore use only photons with energies above 1 \gev.

\begin{shaded} Figure~\ref{all_sensitivity} shows the flux threshold
as a function of source extension for different background levels (1x,
10x, and 100x the nominal background), different spectral indices, and
two different energy bands (1-100 \gev and 10-100 \gev).  The detection
threshold is higher for sources in regions of higher background.  For a
fixed 1-100 \gev or 10-100 \gev flux, the LAT's detection threshold
is only weakly depends upon the spectral index.  This effect is most
pronounced when using only photons with energies above 10 \gev.  Overlaid
on figure~\ref{all_sensitivity} are the LAT detected extended sources
that will be discussed in sections~\ref{validate_known}
and \ref{new_ext_srcs_section}.  The extension thresholds are tabulated
in table~\ref{all_sensitivity_table}.

Finally, figure~\ref{time_sensitivity} shows the LAT's projected detection
threshold to extension after 10 years against 10 times the isotropic
background. This background is representative of the background near
the Galactic plane.  For small extended sources, our detection threshold
improves by a factor larger than the square root of time because at high energies, where we
are most sensitive to extension, the background levels are in the Poisson
instead of Gaussian regime.  For large extended sources, the relevant
background is larger and the improvement is closer to the expected
factor of the square root of time.
\end{shaded}

@@

We test all sources in the 
second Fermi-LAT Catalog (2FGL) for spatial extendedness.
2FGL is a catalog of 1873 sources detected by the LAT.  It included
twelve previously published spatially extended sources but did not
attempt to fit the extension of these sources. Other than these sources,
all 2FGL sources were modeled as point sources and 2FGL did not attempt
to resolve the extension of new sources.

Our analysis setup is closely related to the setup used for the
derivation of the catalog We used the same two years of data from August
4, 2008 to August 1, 2010 and we used the same Pass 7\_V6 (P7\_V6)
Source class event selection and IRFs \citep{lat_on_orbit_psf}.  We used the same models to describe the
background from Galactic diffuse, isotropic, and earth limb emission.
The galactic diffuse emission was scaled by a power-law and the isotropic
component's normalization was left free.

@@

Finally, when analyzing the region, we automatically removed other 2FGL
sources which were within 0.5\deg of the source of interest. This was
done to avoid a concern that extended sources are included in 2FGL as
multiple point sources. These spurious sources could distort the
extension fit.  Instead, when a source is found to be significantly
extended (i.e. $\tsext>16$), we perform the dual localization procedure
(described in section~\ref{dual_localization_method} to compare the
extended source hypothesis to the hypothesis of two independent point
sources. Only sources with $\tsext>\tsinc$ are considered as extended.

@@

% 2FGL J1856.2+0450c = P72Y3047 for
We then inspected the residual \ts, smoothed counts maps, and radial
integral distributions to identify cases in which the extension fit
is clearly influenced by large-scale residuals in the diffuse emission
and hence the extension measurement unreliable.  An example of such a
case is shown in figure~\ref{example_bad_fit}. It shows a Galactic and
isotropic diffuse-emission-subtracted smoothed counts map of this source
using 1-100 \gev photons.  In this region along the Galactic plane,
there appears to be large-scale residual in the diffuse emission. As
a result, 2FGL\,J1856.2+0450c fit to an extension of 1.83\deg and the
result is statistically significant with \tsext=45.4. However, by looking at
the residuals it is clear that this complicated region is not fit well
even though the source's extension is statistically significant. We do
not report sources as extended that fail this inspection.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

For the remaining candidates, 
we took the spatial and spectral model found using \pointlike and
re-determined the spectral 
parameters using
\gtlike.  We used 
the `binned likelihood' mode of \gtlike on a $14\deg\times14\deg$
ROI with a pixel size of 0.03\deg.
We obtained a second measure of \tsext from \gtlike for the extension
and position of the source as determined by \pointlike.
We only considered a source to be significantly extended if $\tsext>16$ with
both \pointlike and \gtlike.

@@

The source 2FGL\,J0823.0-4246 was found 
using 1-100 \gev photons to have an 
extension of $0.37\deg\pm0.03\deg_\stat\pm0.02\deg_\sys$ 
with an extension
significance of $\tsext=46.3$.  The source was fit to a position of
$(l,b)=(260.32\deg,-3.28\deg)$.  Figure~\ref{1FGL_J0823.3-4248} shows a
counts map of this source.  This source is coincident with the one year
catalog source 1FGL\,J0823.3-4248.

@@

\begin{shaded}
This extended source is spatially coincident with the middle-aged
SNR Puppis A.  Puppis A has been studied in detail in radio
(\cite{puppis_a_vla}, and references therein) and  X-ray 
(\cite{rosat_puppis_a,suzaku_puppis_a}, and references therein)
The highest resolution image of the entire source was obtained
by \rosat and contours corresponding to this
image are overlaid on figure~\ref{1FGL_J0823.3-4248} and match the
inferred \gev size \citep{rosat_puppis_a}.
Puppis A's distance was estimated at 2.2 kpc \citep{reynoso_1995,reynoso_2003}
which leads to a 1-100 \gev luminosity of $\approx 3\times 10^{34}$ ergs/\s.

No molecular clouds have been observed directly ajacent to Puppis A 
\citep{co_eastern_puppis_a}.
This is similar to the the LAT detected SNR the Cyngus Loop \citep{cygnus_loop_lat}.
The luminolsity of Puppis A is also smaller than that of 
other SNRs believed to interact with molecular clouds
\citep{w51c,ic443,w44,w28,w49b_lat}.
The \gev emission could therefore be explained by the interaction
of accelerated particles by interstellar matter or fields, similar
to the emission mechanism suggested for the Cygnus Loop SNR
\citep{cygnus_loop_lat}.
\end{shaded}

@@

The source 2FGL\,J1627.0-2425c was found using 1-100 \gev photons to
have an extension of $0.41\deg\pm0.05\deg_\stat\pm0.02\deg_\sys$ with
an extension significance of $\tsext=31.1$.  The best fit position is
$(l,b)=(353.08\deg, 16.78\deg)$.  This source is coincident with the one
year catalog source 1FGL\,J1628.6-2419c.  A counts map showing this source
is seen in figure \ref{1FGL_J1628.6-2419c}.  To test for source confusion,
we also fit the source instead as two point sources. Since $\tsinc=20.0$
is smaller than \tsext, we conclude that this is an extended source.

\begin{shaded}
This source is in a region of remarkably complicated diffuse emission.
Even though it is 16\deg from the Galactic plane, the source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI but also plenty of
dark gas, the kind that we find only by its association with dust emission
\citep{isabelle_dark_gass}. Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and CO
contours are overlaid on figure~\ref{1FGL_J1628.6-2419c} and there is good
spatial correlation with the \gev emission \citep{iras_rho_ophiuci}.
Therefore this source might represent an inadequacy in the diffuse
$\gamma$-ray model
\end{shaded}

@@

The source 2FGL\,J1712.4-3941 was found using 1-100 \gev
photons to have an extension $0.56\deg\pm0.04\deg_\stat\pm0.02\deg_\sys$
with an extension significance of $\tsext=39.6$.  This source was
fit to a position of $(l,b)=(347.25\deg,-0.54\deg)$.  The source
is coincident with the one year catalog source 1FGL\,J1711.7-3944c.
Figure \ref{2FGL_J1712.4-3941} shows a smoothed counts map of this source.

@@

\tablenotetext{a}{Integral 1-100 \mev flux in units of \phflux}

@@

>>>>

@@

% some units
\newcommand{\kev}{\text{Kev}\xspace}
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\newcommand{\km}{\text{km}\xspace}
\newcommand{\phflux}{\ensuremath{\ph\ \cm^{-2}\s^{-1}}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_{\text{ext}}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_{\text{inc}}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}
\newcommand{\likelihood}{\ensuremath{L}\xspace}
\newcommand{\chandra}{\text{{\em Chandra}}\xspace}
\newcommand{\swiftxrt}{\text{{\em Swift}/XRT}\xspace}
\newcommand{\rosat}{\text{{\em ROSAT}}\xspace}
\newcommand{\suzaku}{\text{{\em Suzaku}}\xspace}
\newcommand{\asca}{\text{{\em ASCA}}\xspace}
\newcommand{\einstein}{\text{{\em Einstein}}\xspace}
\newcommand{\xmmnewton}{\text{{\em XMM-Newton}}\xspace}
\newcommand{\fermi}{\textit{Fermi}\xspace}

@@

% the program names
\newcommand{\gtlike}{\ensuremath{\mathtt{gtlike}}\xspace}
\newcommand{\pointlike}{\ensuremath{\mathtt{pointlike}}\xspace}
\newcommand{\gtobssim}{\ensuremath{\mathtt{gtobssim}}\xspace}
\newcommand{\minuit}{\ensuremath{\mathtt{minuit}}\xspace}

\newcommand{\hl}[1]{{\color{red}#1}}

@@

\title{Search for Spatially Extended \fermi-LAT Sources Using Two Years of Flight
Data}
\shorttitle{Search for Extended LAT Sources}

\keywords{
Catalogs;
Fermi Gamma-ray Space Telescope; 
Gamma rays: observations; 
ISM: supernova remnants;
Methods: statistical
pulsar wind nebula
}

@@

\begin{abstract}
We present a new method for analyzing 
the spatial extension of sources with
the Large Area Telescope (LAT), the primary science instrument on the
{\em \fermi Gamma-ray Space Telescope} (\fermi). 
Spatial extension
is an important input to correctly 
associate LAT sources with their counterparts at other
wavelengths. We perform
a series of Monte Carlo studies to validate this tool and calculate
the LAT's detection threshold to detect the extension of 
spatially extended sources.  We then
apply this tool to test all sources in the two year source catalog for
extension. We report the detection of nine spatially extended sources
in addition to the twelve spatially extended sources reported in the
second \fermi-LAT Catalog (2FGL).
\end{abstract}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the {\em \fermi
Gamma-Ray Space Telescope} (\fermi) that been surveying the $\gamma$-ray
sky since 2008 June.  The LAT has a broad energy coverage (20 \mev to $>300$
\gev), wide field of view ($\sim 2.4 \sr$), and large effective area
($\sim 8000 \cm^2$ at $>1 \gev$).

@@

For each of these sources, the extension was determined based on
dedicated analysis of the particular source 
instead of being the outcome of a systematic search of the LAT catalog.
For the systematic scan of all LAT catalog sources presented in this work, we
therefore expect to find additional spatially extended LAT sources, in
particular since many extended sources in the Galactic plane have been
detected at \tev energies using air Cherenkov detectors and \tev and
\gev emission often originates from the same population of high-energy
particles.  A variety of spatially extended sources have been detected
in the Galactic plane at \tev energies. Most prominent was a survey of
the Galactic plane using the High Energy Stereoscopic System (H.E.S.S)
which reported 14 spatially extended sources
 with extensions varying from
$\sim0\fdg1$ to $\sim0\fdg25$ \citep{hess_plane_survey}.
In fact, within
our Galaxy only very few sources (most notably the $\gamma$-ray binaries
LS\,5039 \citep{HESSLS5039}, LS I+61-303 \citep{MAGICLSI, VERITASLSI}
HESS\,J0632+057 \citep{HESS0632}, and the Crab nebula \citep{crab_weekes}
have 
no significant extension detected at \tev energies. Given
these findings, it can be expected that if the source population at \gev
energies has similar properties, that at least some of the \fermi-LAT
sources in our Galaxy should be spatially resolvable at \gev energies.

\hl{
High energy $\gamma$-rays are produced by the decay of $\pi^0$s produced
by hadronic interactions with interstellar matter and by relativistic
electrons due to Inverse Compton (IC) scattering and Bremsstrahlung
radiation.  Many of these spatially extended \tev sources in the
Galactic plane are unidentified. Some of these sources could produce
hadronic cosmic rays up to $\sim 10^{15}$ eV and subsequently radiate
$\gamma$-rays through $\pi^0$ decay \citep{blandford_and_eichler_1987}.
Studying the spectra of these \tev sources at \gev energies may help to
determine the emission mechanisms of these high energy sources.
}

@@

Morphological studies of sources in the \gev energy range using the
LAT are challenging because of the strongly energy-dependent PSF.
Additional complications arise for sources along the Galactic plane due to
systematic uncertainties in the Galactic diffuse emission.  The LAT's
PSF is limited at its lowest detectable energies by multiple scattering in the tracker
and the 68\% containment radius of the PSF is approximately $5\fdg1$
at 100 \mev  when averaged over instrument acceptance and including
photons which convert in either the thick or thin layers of the
tracker. The PSF improves with energy and approaches a limit given
by the granularity of the silicon strips and is $0\fdg14$ at 10 \gev
\citep{on_orbit_calibration}.  However, since most high energy
astrophysical sources have power-law spectra, the better resolution of the higher
energy photons is offset by their low
statistics. Therefore a detailed investigation has been performed in this
work to determine the optimal energy range to measure source extension
(section \ref{extended_source_search_method}).

@@

We estimate the error on the extension of a source by fixing the position
of the source and varying the extension until the likelihood has fallen
by \onehalf, corresponding to a $1\sigma$ error.  This method is
shown schematically in figure~\ref{four_plots_ic443}(a) which shows the
change in \loglikelihood when varying the extension of the SNR IC443.
The localization error is calculated by fixing the extension and spectrum
of the source and fitting to the likelihood function a Gaussian as a function
of position.

@@

We tested two different models for the
surface brightness profile, either a two dimensional Gaussian
\begin{equation}\label{gauss_pdf}
  I(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform disk
\begin{equation}\label{disk_pdf}
  I(x,y)=
  \begin{cases}
    \frac{1}{\pi\sigma^2} & x^2+y^2\le\sigma^2 \\
    0                      & x^2+y^2>\sigma^2
  \end{cases}
\end{equation}
Although these shapes are significantly different, their convolution
with the PSF look similar.  Figure~\ref{compare_disk_gauss}
demonstrates this by showing $I_\text{src}$, the PSF, and the PDF for
a radially symmetric uniform brightness and the Gaussian surface brightness of size 0\fdg5.\footnote{To
allow a valid comparison between Gaussian and disk shaped morphologies,
we define the source size for this test as the radius containing
68\% of the intensity ($\rsixeight$).  $\rsixeight_\text{,Gaussian}=1.51\sigma$
and $\rsixeight_\text{,disk}=0.82\sigma$ where $\sigma$
is defined in equation~\ref{gauss_pdf} and
equation~\ref{disk_pdf} respectively.} This plot shows that the LAT is 
not very sensitive to probing the exact structure of an extended source.
In the search for extended sources, we therefore use only a uniform
disk as our spatial extension shape. We then quote the radius of the
disk edge as the size of the source.

@@

We test the performance of \pointlike fitting the extension of
point-like sources.
\cite{mattox_egret} discuss that the test statistic distribution
for a likelihood ratio test on the existence of a source at
a given position is 
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of equation \ref{ts_ext_distribution} is
due to the null hypothesis (source flux $\Phi=0$) residing
on the edge of parameter space and the model hypothesis
adding a single degree of freedom. It is plausible to
expect a similar distribution of the test statistic
in the test for source extension as the same conditions
apply (with the source flux $\Phi$ replaced by the source radius $r$ and
$r<0$ being unphysical).
To validate this claim, we performed a Monte Carlo
study to calculate empirical distributions for $\tsext$ and
compared them to equation~\ref{ts_ext_distribution}.

@@

For each set of spectral parameters, $\sim30,000$ statistically independent
simulations were performed. For the dimmer spectral models, many of the
simulations left the source undetected ($\ts<25$)
and were discarded.  Table~\ref{ts_ext_num_sims}
shows the different spectral models used in our study as well as the
number of simulations.  The cumulative density of \tsext is plotted in
figure~\ref{ts_ext_mc}. The $\chi^2_1/2$ distribution of
equation~\ref{ts_ext_distribution} is overlaid for comparison.

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0\fdg1$
to $2\fdg0$.  The LAT's flux threshold for a significant detection
of source extension drops quickly with
increasing source size and reaches a minimum around 0\fdg5. 
Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only 1-100\gev photons.
Except for very large ($>1\deg$) sources, our detection threshold is
not substantially improved including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in figure~\ref{four_plots_ic443}(b)
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\sim2.4$, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we therefore use only photons with energies above 1 \gev.

\hl{
Figure~\ref{all_sensitivity} shows the flux threshold
as a function of source extension for different background levels (1x,
10x, and 100x the nominal background), different spectral indices, and
two different energy bands (1-100 \gev and 10-100 \gev).  The detection
threshold is higher for sources in regions of higher background.  For a
fixed 1-100 \gev or 10-100 \gev flux, the LAT's detection threshold
is only weakly depends upon the spectral index.  This effect is most
pronounced when using only photons with energies above 10 \gev.  Overlaid
on figure~\ref{all_sensitivity} are the LAT detected extended sources
that will be discussed in sections~\ref{validate_known}
and \ref{new_ext_srcs_section}.  The extension thresholds are tabulated
in table~\ref{all_sensitivity_table}.

Finally, figure~\ref{time_sensitivity} shows the LAT's projected detection
threshold to extension after 10 years against 10 times the isotropic
background. This background is representative of the background near
the Galactic plane.  For small extended sources, our detection threshold
improves by a factor larger than the square root of time because at high energies, where we
are most sensitive to extension, the background levels are in the Poisson
instead of Gaussian regime.  For large extended sources, the relevant
background is larger and the improvement is closer to the expected
factor of the square root of time.
}

@@

We test all sources in the 
second \fermi-LAT Catalog (2FGL) for spatial extendedness.
2FGL is a catalog of 1873 sources detected by the LAT.  It included
twelve previously published spatially extended sources but did not
attempt to fit the extension of these sources. Other than these sources,
all 2FGL sources were modeled as point sources and 2FGL did not attempt
to resolve the extension of new sources.

Our analysis setup is closely related to the setup used for the
derivation of the catalog We used the same two years of data from 
2008 August 4
to 2010 August 1 and we used the same Pass 7\_V6 (P7\_V6)
Source class event selection and IRFs \citep{lat_on_orbit_psf}.  We used the same models to describe the
background from Galactic diffuse, isotropic, and earth limb emission.
The galactic diffuse emission was scaled by a power-law and the isotropic
component's normalization was left free.

@@

Finally, when analyzing the region, we automatically removed other 2FGL
sources which were within 0\fdg5 of the source of interest. This was
done to avoid a concern that extended sources are included in 2FGL as
multiple point sources. These spurious sources could distort the
extension fit.  Instead, when a source is found to be significantly
extended (i.e. $\tsext>16$), we perform the dual localization procedure
(described in section~\ref{dual_localization_method} to compare the
extended source hypothesis to the hypothesis of two independent point
sources. Only sources with $\tsext>\tsinc$ are considered as extended.

@@

% 2FGL J1856.2+0450c = P72Y3047 for
We then inspected the residual \ts, smoothed counts maps, and radial
integral distributions to identify cases in which the extension fit
is clearly influenced by large-scale residuals in the diffuse emission
and hence the extension measurement unreliable.  An example of such a
case is shown in figure~\ref{example_bad_fit}. It shows a Galactic and
isotropic diffuse-emission-subtracted smoothed counts map of this source
using 1-100 \gev photons.  In this region along the Galactic plane,
there appears to be large-scale residual in the diffuse emission. As
a result, 2FGL\,J1856.2+0450c fit to an extension of 1\fdg83 and the
result is statistically significant with \tsext=45.4. However, by looking at
the residuals it is clear that this complicated region is not fit well
even though the source's extension is statistically significant. We do
not report sources as extended that fail this inspection.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

For the remaining candidates, 
we took the spatial and spectral model found using \pointlike and
re-determined the spectral 
parameters using
\gtlike.  We used 
the `binned likelihood' mode of \gtlike on a $14\deg\times14\deg$
ROI with a pixel size of 0\fdg03.
We obtained a second measure of \tsext from \gtlike for the extension
and position of the source as determined by \pointlike.
We only considered a source to be significantly extended if $\tsext>16$ with
both \pointlike and \gtlike.

@@

The source 2FGL\,J0823.0-4246 was found 
using 1-100 \gev photons to have an 
extension of $0\fdg37\pm0\fdg03_\stat\pm0\fdg02_\sys$ 
with an extension
significance of $\tsext=46.3$.  The source was fit to a position of
$(l,b)=(260\fdg32,-3\fdg28)$.  
This source is coincident with the one year
catalog source 1FGL\,J0823.3-4248.
\hl{
Figure~\ref{1FGL_J0823.3-4248} shows a
counts map of this source
and figure~\ref{snr_seds}(a) shows the LAT spectrum of it.
}

@@

\hl{1
This extended source is spatially coincident with the middle-aged
SNR Puppis A.  Puppis A has been studied in detail in radio
(\cite{puppis_a_vla}, and references therein) and  X-ray 
(\cite{rosat_puppis_a,suzaku_puppis_a}, and references therein)
The highest resolution image of the entire source was obtained
by \rosat and contours corresponding to this
image are overlaid on figure~\ref{1FGL_J0823.3-4248} and match the
inferred \gev size \citep{rosat_puppis_a}.
Puppis A's distance was estimated at 2.2 kpc \citep{reynoso_1995,reynoso_2003}
which leads to a 1-100 \gev luminosity of $\sim 3\times 10^{34}$ ergs/\s.

No molecular clouds have been observed directly ajacent to Puppis A 
\citep{co_eastern_puppis_a}.
This is similar to the the LAT detected SNR the Cyngus Loop \citep{cygnus_loop_lat}.
The luminolsity of Puppis A is also smaller than that of 
other SNRs believed to interact with molecular clouds
\citep{w51c,ic443,w44,w28,w49b_lat}.
The \gev emission could therefore be explained by the interaction
of accelerated particles by interstellar matter or fields, similar
to the emission mechanism suggested for the Cygnus Loop SNR
\citep{cygnus_loop_lat}.
}

@@

The source 2FGL\,J1627.0-2425c was found using 1-100 \gev photons to
have an extension of $0\fdg41\pm0\fdg05_\stat\pm0\fdg02_\sys$ with
an extension significance of $\tsext=31.1$.  The best fit position is
$(l,b)=(353\fdg08, 16\fdg78)$.  This source is coincident with the one
year catalog source 1FGL\,J1628.6-2419c.  A counts map showing this source
is seen in figure \ref{1FGL_J1628.6-2419c}.  To test for source confusion,
we also fit the source instead as two point sources. Since $\tsinc=20.0$
is smaller than \tsext, we conclude that this is an extended source.

\hl{
This source is in a region of remarkably complicated diffuse emission.
Even though it is $16\deg$ from the Galactic plane, the source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI but also plenty of
dark gas, the kind that we find only by its association with dust emission
\citep{isabelle_dark_gass}. Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and 
${}^{12}\text{CO}$ ($J=1\rightarrow 0$)
contours are overlaid on figure~\ref{1FGL_J1628.6-2419c}. There is good
spatial correlation with the \gev emission \citep{iras_rho_ophiuci,co_rho_ophiuci}
so this source might represent an inadequacy in the diffuse
$\gamma$-ray model.
}

@@

The source 2FGL\,J1712.4-3941 was found using 1-100 \gev
photons to have an extension $0\fdg56\pm0\fdg04_\stat\pm0\fdg02_\sys$
with an extension significance of $\tsext=39.6$.  This source was
fit to a position of $(l,b)=(347\fdg25,-0\fdg54)$.  The source
is coincident with the one year catalog source 1FGL\,J1711.7-3944c.
Figure \ref{2FGL_J1712.4-3941} shows a smoothed counts map of this source.

@@

\tablenotetext{(a)}{Integral 1-100 \mev flux.}

@@

\tablecolumns{22}
